More equivalences for Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-10-11 2016-Weak-FH-Bisim/WeakBisim.tex
small changes : vis is v and overline is set

%%\smallskip\noindent
%\paragraph*{Indexed sets.}
%We extensively use indexed structures
%over some countable indexed sets, which are equivalent to mappings over
%the countable set. % . The indexes will usually be
%% integers, bounded or not. Such an indexed family is
%%denoted
%%follows:
%$a_i^{i\in I}$
%%, or equivalently  $(i\mapsto a_i)^{i\in I}$
%denotes a family of elements $a_i$ indexed over the
%set $I$. % Such a family
%% is equivalent to the mapping $(i\mapsto a_i)^{i\in I}$.
%% To specify the set over which the structure is indexed,
%% indexed structures are always denoted with an exponent of the form $i\in I$
%% (arithmetic only appears in the indexes if necessary).
%$a_i^{i\in I}$ defines both $I$ the set over which the family is
%indexed (called \emph{range}), and $a_i$ the elements of the family.
%E.g., $a^{i\in\{3\}}$ is the mapping with a single entry $a$ at index
%$3$ ; abbreviated $(3\mapsto a)$ in the following.
%When this is not
%ambiguous, we shall use notations for sets, and typically write
%``indexed set over I'' when formally we should speak of multisets, and
%write $x\in a_i^{i\in I}$ to mean $\exists i\in I.\, x=a_i$.  An empty
%family is denoted $\emptyset$. We
%denote classically with an overline -- $\overline{a}$  -- a family when the indexing set 
%is
%not meaningful.  $\uplus$ is the disjoint union on
%indexed sets.

@@

\begin{definition}[Sorts, Holes, Leaves of pNets]\label{def-sortholeleave}
  \begin{itemize}
  \item The sort of a pNet is its signature, i.e. the set of actions it can
perform, where each action signature is an action 
label plus the arity of the action. In the definition of sorts, we do not need to 
distinguish
input variables and we remove the
\emph{input marker} (?) of variables.
\[
\begin{array}{l}
\Sortop(\mylangle S,s_0, \to\myrangle) = \{\alpha\subst{x \gets ?x| 
x\in\symb{iv}(\alpha)}|s \xrightarrow{\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in
    J}\rangle} s'\in \to \} \\
\Sortop(\mylangle \overline{\pNet}\!, %\pNet_i^{i\in I}, 
\overline{\Sort},
\overline{\symb{SV}}\myrangle)
=\{\alpha' |\, \SV{\overline{\alpha}}{\alpha'}{e_b}\in\set{\symb{SV}}\}
\end{array}
\]

\item
The set of holes $\Holes(\pNet)$ of a pNet is the indexes of the holes of the pNet 
itself plus the indexes of all the holes of its subnets.
It is defined inductively (we suppose those indexes 
disjoints):
  \[\begin{array}{l}
\Holes(\mylangle S,s_0, \to\myrangle) \!=\! \emptyset\\
\Holes(\mylangle \pNet_i^{i\in I}\!,\overline{\Sort}, \overline{\symb{SV}}\myrangle) 
=J\uplus{\displaystyle \bigcup_{i\in 
I}\Holes(\pNet_i)}\\
\forall i\in I.\, \Holes(\pNet_i)\cap J=\emptyset\\
\forall i_1,i_2\in I.\,i_1\neq i_2\Rightarrow  \Holes(\pNet_{i_1})\cap\Holes(\pNet_{i_2})=\emptyset
\end{array}\]
\item
The set of leaves of a pNet is the set of all pLTSs occurring in the structure, as an 
indexed family of the form $\Leaves(\pNet)= \mylangle \pNet_i \myrangle^{i \in L}$.
\[\begin{array}{l}
\Leaves(\mylangle S,s_0, \to\myrangle) \!=\!\emptyset\\%\{\mylangle S,s_0, \to\myrangle\}\\
\Leaves(\mylangle \pNet_i^{i\in I}\!,%Sort_j^{j\in J}
\overline{\Sort}\!, \overline{\symb{SV}}\myrangle) = {\displaystyle \biguplus_{i\in 
I}\Leaves(\pNet_i)\uplus\{i\mapsto \pNet_i|\pNet_i \text{ is a }\pLTS\}}
\end{array}\]
\end{itemize}

@@

\begin{definition}[pNet composition]
	An open pNet: $\pNet = \mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, 
	\overline{\symb{SV}}\myrangle$
 can be (partially) filled by providing  a pNet $\pNetQ$ of the
	right sort to fill one of  its holes.	
	Suppose $j_0\in J$:
	%% \[\mylangle \pNet_i^{i\in I},S_j^{j\in J}, \symb{SV}_k^{k\in
	%%   K}\myrangle\left[(\pNet'_i)^{i\in L}\right]= \mylangle \pNet_i^{i\in
	%%   I}\left[(\pNet'_i)^{i\in L}\right]\uplus(\pNet'_i)^{i\in J\cap L},S_j^{j\in 
	%%J\setminus L},
	%% \symb{SV}_k^{k\in K}\myrangle
	%% \]
	\[\pNet\left[\pNetQ\right]_{j_0}= \mylangle 
	\pNet_i^{i\in I}\uplus\{j_0\mapsto \pNetQ\},\Sort_j^{j\in J\setminus \{j_0\}},
	\overline{\symb{SV}}\myrangle
	\]
\end{definition}

@@

\end{mathpar}
 there exist open transitions $OT_x^{x\in X} \subseteq \mathcal{T}_2$:
 \begin{mathpar}
%    \left( fresh \ \overline{\alpha_i}, \overline{b_j}, v_x.\ \
    \openrule
         {
           \beta_{j x}^{j\in J_{x}}, \Pred_{OT_x},\Post_{OT_x}}
         {t \OTarrow {\alpha_x} t_x}
%         \right)
\end{mathpar}
 such that  $\forall x, J'=J_{x}, \exists \Pred_{s',t_x}. (s',t_x|\Pred_{s',t_x})\in 
 \mathcal{R}$; 
 and  \\
 $\Pred_{s,t} \land \Pred_{OT}\\
\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
   \left( \forall j. \beta_j=\beta_{jx}  \Rightarrow \Pred_{OT_x}
     \land \alpha\!=\!\alpha_x \land  
     \Pred_{s',t_x}\subst{\Post_{OT}}\subst{\Post_{OT_x}}\right)$
     %     \symb{Subst}(\Pred_{target_x}, \Post_{OT} o \Post_{OT_x})
     %     \right)$.

@@

\paragraph{Predicates:}
%Let
%$\mylangle\overline{\pNet},\overline{\Sort},\overline{\symb{SV}}\myrangle$
%be a pNet. 
Consider a synchronisation vector $\symb{SV}$. We 
define a
predicate $\Predsv$ relating
the actions of the involved sub-pNets and the resulting actions. This predicate verifies:
\[\begin{array}{l}\Predsv \Big(\SV{{(\alpha'_i)}^{i\in I}, {(\beta'_j)}^{j\in J}} 
{\alpha'} 
{e_b}, \alpha_i^{i\in I}, \beta_j^{j\in J}, \alpha\Big)\Leftrightarrow \\~\qquad\qquad%\bigg(
%
%\exists {(\alpha'_i)}^{i\in I},
%{(\beta'_j)}^{j\in J},v'.\, SV=
%\\~~\land
\forall i\in I.\, \alpha_i=\alpha'_i\land \forall j \in J.\, \beta_j=\beta'_j \land 
\alpha=\alpha' 
\land e_b
\end{array} 
%\bigg)
\]

@@

\begin{definition}[Operational semantics of open pNets]
	\label{def:operationalSemantics}
	The semantics of a pNet $\pNet$ is an open automaton $A\!= 
	<\!\!Holes(\pNet),States(\pNet),InitState(\pNet),\mathcal{T}\!\!>$ where $\mathcal{T}$   is the smallest set of open transitions such that $\mathcal{T}=\{OT\,|\,\pNet \models OT \}$ and	$\pNet \models OT$	is defined by the following  rules:
%	\begin{itemize}
%		\item $J$ is the set of holes: $Holes(p)= J$. 
		%  \item $\overline{L}^L = Leaves(p), \overline{H}^J = Holes(p)$
%		\item ${\mathcal{S}} = States(p)$ and $s_0 = InitState(p)$
%		\item $\mathcal{T}$ is the smallest set of open transitions		satisfying the rules below:
%	\end{itemize}

@@

	\noindent
\begin{mathpar}
    \mprset {vskip=.8ex}
\inferrule
    {
\Leaves(\mylangle \overline{\pNet}, \overline{\Sort}, \symb{SV}_k^{k\in 
    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \qquad  	
k\!\in\! K \qquad SV_k \!=\! \SV{(\alpha'_m)^{m \in I_1\uplus I_2\uplus J}}{\alpha'}{e_b} 
\\
\\     	
	\forall m\!\!\in\!\! I_1. {\pNet_m 
	\models\openrule
    	{
    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}} }	
  \qquad
\forall m\!\!\in\!\! I_2.		{ \pNet_m 
    	 \models
    	\openrule
    	{\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {\alpha_m}
    		\ostate{s_m'}} }\\\\
     J' = \biguplus_{m\in I_1}\!\! J_m \uplus J 	\\
    	\Pred = \bigwedge_{m\in I_1\uplus I_2}\!\! \Pred_m \land
    	\Predsv(SV_k,\alpha_m^{m\in I_1\uplus I_2},\beta_j^{j\in J},\alpha)\\ 
    		I' = \biguplus_{m\in I_1}\!\! I_m \uplus I_2
    	\\\forall i\in	L\backslash I'.\,s'_i=s_i \\
    \fresh(\alpha'_m,\alpha',\beta_j,\alpha) 
    }
    {\mylangle \overline{\pNet}, \overline{\Sort}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\openrule
    		{
    		{\beta_j}^{j\in J^\prime}, \Pred,  \biguplus_{m\in I_1\uplus I_2} 
    		\Post_m}
    		{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
    			\ostate{(s_i^\prime)^{i\in L}}}
    	}
    }
\end{mathpar}    
	\medskip
%        \TODO{may be explain how $\Pred(SV,a_i^{i\in I_k},b_j^{j\in
%            J_k},v)$ is built ? You mean more than what is written on previous page????}
	%%    \TODO{I have tentatively added the sort constraint on hole actions, that was
	%%not included in the first version... I'm unsure whether this is the best place to
	%%include it, because it may change the decidability conditions on predicates}
\end{definition}
        	A key to understand this rule is that the open transitions are
	expressed in terms of the leaves and holes of the pNet structure,
	i.e. a flatten view of the pNet: e.g. $L$ is the index set of the
	Leaves, $L_m$ the index set of the leaves of one subnet indexed $m$, so all $L_m$
	are disjoint subsets of $L$. Thus the states in the open transitions,
	at each level, are tuples including states of all the
	leaves of the pNet, not only those involved in the chosen
	synchronisation vector.

@@

\begin{theorem}[Congruence]\label{thm-congr-eq}
	Consider an open pNet:
	$\pNet = \mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, 
	\overline{\symb{SV}}\myrangle$.
	Let $j_0\in J$ be a hole. Let $\pNetQ$ and $\pNetQ'$ be two FH-bisimilar pNets such that 
	$\Sortop(\pNetQ)=\Sortop(\pNetQ')=\Sort_{j_0}$\footnote{Note that $\Sortop(\pNetQ)=\Sortop(\pNetQ')$ is 
	ensured by 
	strong bisimilarity.}. Then 
	$\pNet[\pNetQ]_{j_0}$ and 
	$\pNet[\pNetQ']_{j_0}$ are FH-bisimilar.
\end{theorem}

\begin{theorem}[Context equivalence]\label{thm-ctxt-eq}
	Consider two FH-bisimilar open pNets:
	$\pNet = \mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, 
	\overline{\symb{SV}}\myrangle$ and 	$\pNet' = \mylangle {\pNet'}_i^{i\in I}, 
	\Sort_j^{j\in 
	J}, 	\overline{\symb{SV'}}\myrangle$ 
	(recall they must have the same holes to be bisimilar).
	Let $j_0\in J$ be a hole, and $Q$ be a pNet such that $\Sortop(Q)=\Sort_{j_0}$. Then 
	$\pNet[Q]_{j_0}$ and 
	$\pNet'[Q]_{j_0}$ are FH-bisimilar.
\end{theorem}

@@

Intuition: tau is non observable if the automaton allows any tau transition from holes, and the global transition resulting from a tau in a hole is a tau transition not changing the action state.
\TODO{Id(s) is the identity function on variables of state s}
\begin{definition}[Non-observability of silent actions]
An open automaton $A = <J,\mathcal{S},s_0,\mathcal{T}>$ \emph{cannot observe silent actions} if and only if for all $j$ in $J$ and $s$ in $\mathcal{S}$ we have
\begin{enumerate}
\item
\[ \openrule
         {
           (j\mapsto\tau),\True,\Id(s)}
         {s \OTarrow {\tau} s}
         \in \mathcal{T}
\]
and 
\item for all $\alpha$, $j$, $s$, $\Post$ and $s'$ such that
\[ \openrule
         {
           (j\mapsto\tau),\Pred,\Post}
         {s \OTarrow {\alpha} s'}
         \in \mathcal{T} \] we have \[ \alpha=\tau\land s=s'\land \Pred=\True\land\Post=\Id(s)
\]
\end{enumerate}
\end{definition}

@@

We define $\vis{\overline\beta}$ as the mapping $\overline\beta$  without  invisible actions but where each element is now a list (if length 1).\\ $\vis{\beta_i^{i\in I}} = [\beta_i]^{i\in I'}$ s.t. $I'=\left\{i| i\in I \land \beta_i\neq \tau\right\}$.

@@

%\noindent{\bf Invisible $\tau$ transitions:}
\begin{mathpar}
 \openrule
         {
           \emptyset,\symb{True},\Id(s)}
         {s \OTWeakarrow {\tau} s} \in \WT  \qquad \WTUn
 \end{mathpar}
and
\begin{mathpar}
  \mprset {vskip=.5ex}
\inferrule{
 \openrule
         {
           \overline{\beta},\Pred,\Post}
         {s \OTarrow {\alpha} s'} \in \mathcal{T}
} 
{ \openrule
         {
           \vis{\overline{\beta}}\!,\Pred,\Post
				 } {s \OTWeakarrow {\alpha} s'} \in \WT
}\qquad \WTDeux
%\inferrule{
% \openrule
%         {
%           \overline{\beta},\Pred,\Post}
%         {s \OTarrow {\tau} s'} \in \mathcal{T}
%\\
% \openrule
%         {
%           \overline{\gamma},\Pred\,',({x_k\gets e_k})^{k\in K}   }
%         {s' \OTWeakarrow {\tau} s''} \in \WT
%}
%{ \openrule
%         {
%           \vis{\overline{\beta}}\dotcup\overline{\gamma},\Pred\land\Pred\,'\subst{\Post\,},
%				({x_k\gets (e_k\subst{\Post\,})})^{k\in K} } 
%         {s \OTWeakarrow {\tau} s''} \in \WT
%}
 \end{mathpar}
 and
%\noindent{\bf Other transitions:}
\begin{mathpar}
  \mprset {vskip=.5ex}
\inferrule {\openrule
         {
           \overline{\gamma_1},\Pred_1,\Post_1   }
         {s \OTWeakarrow {\tau} s_1} \in \WT
\qquad
\openrule
         {
           \overline{\gamma_2},\Pred_2,\Post_2  }
         {s_1 \OTWeakarrow {\alpha} s_2} \in \WT
\qquad
\openrule
         {
           \overline{\gamma_3},\Pred_3,({x_k\gets e_k})^{k\in K}    }
         {s_2 \OTWeakarrow {\tau} s'} \in\WT
\\
\Pred=\Pred_1\land\Pred_2\subst{\Post_1}\land (\Pred_3\subst{\Post_2})\subst{\Post_1}
}
{
\openrule
         {
           \overline{\gamma_1}\dotcup \overline{\gamma_2}\dotcup\overline{\gamma_3},
		\Pred,
				({x_k\gets (e_k\subst {\Post_2})\subst{\Post_1}})^{k\in K} } 
         {s \OTWeakarrow {\alpha} s'} \in\WT
} \WTTrois
\end{mathpar}

@@

\subsection{Weak bisimulation for open pNets}
\begin{definition}[Non-observability of silent actions for pNets]~\\
A pNet $\mylangle \pNet_i^{i\in I} , \Sort_j^{j\in J}, \overline{\symb{SV}}\myrangle$
 \emph{cannot observe silent actions} if it verifies:\\ $\forall i\in I\uplus J.\, \SV{(i\mapsto \tau)}{\tau}{\True}\in \overline{\symb{SV}}$ and $\forall SV_k\in \overline{\symb{SV}}, \forall i,j\in I_k\uplus J_k.\,i\neq j \Rightarrow \alpha'_k\neq\tau$
\end{definition}

@@

%%    \left( fresh \ \overline{a_i}, \overline{b_j}, v_x.\ \

@@

       	\begin{mathpar} 
       	%    \left( fresh \ \overline{a_i}, \overline{b_j}, v_x.\ \
       	\openrule
       	{
       		\beta_{j p x}^{j\in J_{px}}, \Pred_{OT_{px}},\Post_{OT_{px}}}
       	{t_p \OTarrow {\alpha_{px}} t_{p x}}\qquad (*)
       	%         \right)
       	\end{mathpar}

>>>>

%%\smallskip\noindent
%\paragraph*{Indexed sets.}
%We extensively use indexed structures
%over some countable indexed sets, which are equivalent to mappings over
%the countable set. % . The indexes will usually be
%% integers, bounded or not. Such an indexed family is
%%denoted
%%follows:
%$a_i^{i\in I}$
%%, or equivalently  $(i\mapsto a_i)^{i\in I}$
%denotes a family of elements $a_i$ indexed over the
%set $I$. % Such a family
%% is equivalent to the mapping $(i\mapsto a_i)^{i\in I}$.
%% To specify the set over which the structure is indexed,
%% indexed structures are always denoted with an exponent of the form $i\in I$
%% (arithmetic only appears in the indexes if necessary).
%$a_i^{i\in I}$ defines both $I$ the set over which the family is
%indexed (called \emph{range}), and $a_i$ the elements of the family.
%E.g., $a^{i\in\{3\}}$ is the mapping with a single entry $a$ at index
%$3$ ; abbreviated $(3\mapsto a)$ in the following.
%When this is not
%ambiguous, we shall use notations for sets, and typically write
%``indexed set over I'' when formally we should speak of multisets, and
%write $x\in a_i^{i\in I}$ to mean $\exists i\in I.\, x=a_i$.  An empty
%family is denoted $\emptyset$. We
%denote classically with an overline -- $\set{a}$  -- a family when the indexing set 
%is
%not meaningful.  $\uplus$ is the disjoint union on
%indexed sets.

@@

\begin{definition}[Sorts, Holes, Leaves of pNets]\label{def-sortholeleave}
  \begin{itemize}
  \item The sort of a pNet is its signature, i.e. the set of actions it can
perform, where each action signature is an action 
label plus the arity of the action. In the definition of sorts, we do not need to 
distinguish
input variables and we remove the
\emph{input marker} (?) of variables.
\[
\begin{array}{l}
\Sortop(\mylangle S,s_0, \to\myrangle) = \{\alpha\subst{x \gets ?x| 
x\in\symb{iv}(\alpha)}|s \xrightarrow{\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in
    J}\rangle} s'\in \to \} \\
\Sortop(\mylangle \set{\pNet}\!, %\pNet_i^{i\in I}, 
\set{\Sort},
\set{\symb{SV}}\myrangle)
=\{\alpha' |\, \SV{\set{\alpha}}{\alpha'}{e_b}\in\set{\symb{SV}}\}
\end{array}
\]

\item
The set of holes $\Holes(\pNet)$ of a pNet is the indexes of the holes of the pNet 
itself plus the indexes of all the holes of its subnets.
It is defined inductively (we suppose those indexes 
disjoints):
  \[\begin{array}{l}
\Holes(\mylangle S,s_0, \to\myrangle) \!=\! \emptyset\\
\Holes(\mylangle \pNet_i^{i\in I}\!,\set{\Sort}, \set{\symb{SV}}\myrangle) 
=J\uplus{\displaystyle \bigcup_{i\in 
I}\Holes(\pNet_i)}\\
\forall i\in I.\, \Holes(\pNet_i)\cap J=\emptyset\\
\forall i_1,i_2\in I.\,i_1\neq i_2\Rightarrow  \Holes(\pNet_{i_1})\cap\Holes(\pNet_{i_2})=\emptyset
\end{array}\]
\item
The set of leaves of a pNet is the set of all pLTSs occurring in the structure, as an 
indexed family of the form $\Leaves(\pNet)= \mylangle \pNet_i \myrangle^{i \in L}$.
\[\begin{array}{l}
\Leaves(\mylangle S,s_0, \to\myrangle) \!=\!\emptyset\\%\{\mylangle S,s_0, \to\myrangle\}\\
\Leaves(\mylangle \pNet_i^{i\in I}\!,%Sort_j^{j\in J}
\set{\Sort}\!, \set{\symb{SV}}\myrangle) = {\displaystyle \biguplus_{i\in 
I}\Leaves(\pNet_i)\uplus\{i\mapsto \pNet_i|\pNet_i \text{ is a }\pLTS\}}
\end{array}\]
\end{itemize}

@@

\begin{definition}[pNet composition]
	An open pNet: $\pNet = \mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, 
	\set{\symb{SV}}\myrangle$
 can be (partially) filled by providing  a pNet $\pNetQ$ of the
	right sort to fill one of  its holes.	
	Suppose $j_0\in J$:
	%% \[\mylangle \pNet_i^{i\in I},S_j^{j\in J}, \symb{SV}_k^{k\in
	%%   K}\myrangle\left[(\pNet'_i)^{i\in L}\right]= \mylangle \pNet_i^{i\in
	%%   I}\left[(\pNet'_i)^{i\in L}\right]\uplus(\pNet'_i)^{i\in J\cap L},S_j^{j\in 
	%%J\setminus L},
	%% \symb{SV}_k^{k\in K}\myrangle
	%% \]
	\[\pNet\left[\pNetQ\right]_{j_0}= \mylangle 
	\pNet_i^{i\in I}\uplus\{j_0\mapsto \pNetQ\},\Sort_j^{j\in J\setminus \{j_0\}},
	\set{\symb{SV}}\myrangle
	\]
\end{definition}

@@

\end{mathpar}
 there exist open transitions $OT_x^{x\in X} \subseteq \mathcal{T}_2$:
 \begin{mathpar}
%    \left( fresh \ \set{\alpha_i}, \set{b_j}, v_x.\ \
    \openrule
         {
           \beta_{j x}^{j\in J_{x}}, \Pred_{OT_x},\Post_{OT_x}}
         {t \OTarrow {\alpha_x} t_x}
%         \right)
\end{mathpar}
 such that  $\forall x, J'=J_{x}, \exists \Pred_{s',t_x}. (s',t_x|\Pred_{s',t_x})\in 
 \mathcal{R}$; 
 and  \\
 $\Pred_{s,t} \land \Pred_{OT}\\
\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
   \left( \forall j. \beta_j=\beta_{jx}  \Rightarrow \Pred_{OT_x}
     \land \alpha\!=\!\alpha_x \land  
     \Pred_{s',t_x}\subst{\Post_{OT}}\subst{\Post_{OT_x}}\right)$
     %     \symb{Subst}(\Pred_{target_x}, \Post_{OT} o \Post_{OT_x})
     %     \right)$.

@@

\paragraph{Predicates:}
%Let
%$\mylangle\set{\pNet},\set{\Sort},\set{\symb{SV}}\myrangle$
%be a pNet. 
Consider a synchronisation vector $\symb{SV}$. We 
define a
predicate $\Predsv$ relating
the actions of the involved sub-pNets and the resulting actions. This predicate verifies:
\[\begin{array}{l}\Predsv \Big(\SV{{(\alpha'_i)}^{i\in I}, {(\beta'_j)}^{j\in J}} 
{\alpha'} 
{e_b}, \alpha_i^{i\in I}, \beta_j^{j\in J}, \alpha\Big)\Leftrightarrow \\~\qquad\qquad%\bigg(
%
%\exists {(\alpha'_i)}^{i\in I},
%{(\beta'_j)}^{j\in J},v'.\, SV=
%\\~~\land
\forall i\in I.\, \alpha_i=\alpha'_i\land \forall j \in J.\, \beta_j=\beta'_j \land 
\alpha=\alpha' 
\land e_b
\end{array} 
%\bigg)
\]

@@

\begin{definition}[Operational semantics of open pNets]
	\label{def:operationalSemantics}
	The semantics of a pNet $\pNet$ is an open automaton $A\!= 
	<\!\!Holes(\pNet),States(\pNet),InitState(\pNet),\mathcal{T}\!\!>$ where $\mathcal{T}$   is the smallest set of open transitions such that $\mathcal{T}=\{OT\,|\,\pNet \models OT \}$ and	$\pNet \models OT$	is defined by the following  rules:
%	\begin{itemize}
%		\item $J$ is the set of holes: $Holes(p)= J$. 
		%  \item $\set{L}^L = Leaves(p), \set{H}^J = Holes(p)$
%		\item ${\mathcal{S}} = States(p)$ and $s_0 = InitState(p)$
%		\item $\mathcal{T}$ is the smallest set of open transitions		satisfying the rules below:
%	\end{itemize}

@@

	\noindent
\begin{mathpar}
    \mprset {vskip=.8ex}
\inferrule
    {
\Leaves(\mylangle \set{\pNet}, \set{\Sort}, \symb{SV}_k^{k\in 
    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \qquad  	
k\!\in\! K \qquad SV_k \!=\! \SV{(\alpha'_m)^{m \in I_1\uplus I_2\uplus J}}{\alpha'}{e_b} 
\\
\\     	
	\forall m\!\!\in\!\! I_1. {\pNet_m 
	\models\openrule
    	{
    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}} }	
  \qquad
\forall m\!\!\in\!\! I_2.		{ \pNet_m 
    	 \models
    	\openrule
    	{\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {\alpha_m}
    		\ostate{s_m'}} }\\\\
     J' = \biguplus_{m\in I_1}\!\! J_m \uplus J 	\\
    	\Pred = \bigwedge_{m\in I_1\uplus I_2}\!\! \Pred_m \land
    	\Predsv(SV_k,\alpha_m^{m\in I_1\uplus I_2},\beta_j^{j\in J},\alpha)\\ 
    		I' = \biguplus_{m\in I_1}\!\! I_m \uplus I_2
    	\\\forall i\in	L\backslash I'.\,s'_i=s_i \\
    \fresh(\alpha'_m,\alpha',\beta_j,\alpha) 
    }
    {\mylangle \set{\pNet}, \set{\Sort}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\openrule
    		{
    		{\beta_j}^{j\in J^\prime}, \Pred,  \biguplus_{m\in I_1\uplus I_2} 
    		\Post_m}
    		{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
    			\ostate{(s_i^\prime)^{i\in L}}}
    	}
    }
\end{mathpar}    
	\medskip
%        \TODO{may be explain how $\Pred(SV,a_i^{i\in I_k},b_j^{j\in
%            J_k},v)$ is built ? You mean more than what is written on previous page????}
	%%    \TODO{I have tentatively added the sort constraint on hole actions, that was
	%%not included in the first version... I'm unsure whether this is the best place to
	%%include it, because it may change the decidability conditions on predicates}
\end{definition}
        	A key to understand this rule is that the open transitions are
	expressed in terms of the leaves and holes of the pNet structure,
	i.e. a flatten view of the pNet: e.g. $L$ is the index set of the
	Leaves, $L_m$ the index set of the leaves of one subnet indexed $m$, so all $L_m$
	are disjoint subsets of $L$. Thus the states in the open transitions,
	at each level, are tuples including states of all the
	leaves of the pNet, not only those involved in the chosen
	synchronisation vector.

@@

\begin{theorem}[Congruence]\label{thm-congr-eq}
	Consider an open pNet:
	$\pNet = \mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, 
	\set{\symb{SV}}\myrangle$.
	Let $j_0\in J$ be a hole. Let $\pNetQ$ and $\pNetQ'$ be two FH-bisimilar pNets such that 
	$\Sortop(\pNetQ)=\Sortop(\pNetQ')=\Sort_{j_0}$\footnote{Note that $\Sortop(\pNetQ)=\Sortop(\pNetQ')$ is 
	ensured by 
	strong bisimilarity.}. Then 
	$\pNet[\pNetQ]_{j_0}$ and 
	$\pNet[\pNetQ']_{j_0}$ are FH-bisimilar.
\end{theorem}

\begin{theorem}[Context equivalence]\label{thm-ctxt-eq}
	Consider two FH-bisimilar open pNets:
	$\pNet = \mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, 
	\set{\symb{SV}}\myrangle$ and 	$\pNet' = \mylangle {\pNet'}_i^{i\in I}, 
	\Sort_j^{j\in 
	J}, 	\set{\symb{SV'}}\myrangle$ 
	(recall they must have the same holes to be bisimilar).
	Let $j_0\in J$ be a hole, and $Q$ be a pNet such that $\Sortop(Q)=\Sort_{j_0}$. Then 
	$\pNet[Q]_{j_0}$ and 
	$\pNet'[Q]_{j_0}$ are FH-bisimilar.
\end{theorem}

@@

Intuition: tau is non observable if the automaton allows any tau transition from holes, and the global transition resulting from a tau in a hole is a tau transition not changing the action state.
We define $\Id(s)$ as the identity function on variables of state s.
\begin{definition}[Non-observability of silent actions]
An open automaton $A = <J,\mathcal{S},s_0,\mathcal{T}>$ \emph{cannot observe silent actions} if and only if for all $j$ in $J$ and $s$ in $\mathcal{S}$ we have
\begin{enumerate}
\item
\[ \openrule
         {
           (j\mapsto\tau),\True,\Id(s)}
         {s \OTarrow {\tau} s}
         \in \mathcal{T}
\]
and 
\item for all $\alpha$, $j$, $s$, $\Post$ and $s'$ such that
\[ \openrule
         {
           (j\mapsto\tau),\Pred,\Post}
         {s \OTarrow {\alpha} s'}
         \in \mathcal{T} \] we have \[ \alpha=\tau\land s=s'\land \Pred=\True\land\Post=\Id(s)
\]
\TODO{question : it looks like we can have 2 holes doing a tau and observe this ; I belive this is not safe (LUDO)}
\end{enumerate}
\end{definition}
The first statement of the definition state that the open automaton must allow a hole to do a silent action at any time, and must not observe it, i.e. cannot change its internal state because a hole did a $\tau$ transition. The second statement ensures that there cannot be in the open automaton other transitions that would be able to observe a silent action from a hole.

@@

We define $\vis{\set\beta}$ as the mapping $\set\beta$  without  invisible actions but where each element is now a list (if length 1).\\ $\vis{\beta_i^{i\in I}} = [\beta_i]^{i\in I'}$ s.t. $I'=\left\{i| i\in I \land \beta_i\neq \tau\right\}$.

@@

%\noindent{\bf Invisible $\tau$ transitions:}
\begin{mathpar}
 \openrule
         {
           \emptyset,\symb{True},\Id(s)}
         {s \OTWeakarrow {\tau} s} \in \WT  \qquad \WTUn
 \end{mathpar}
and
\begin{mathpar}
  \mprset {vskip=.5ex}
\inferrule{
 \openrule
         {
           \set{\beta},\Pred,\Post}
         {s \OTarrow {\alpha} s'} \in \mathcal{T}
} 
{ \openrule
         {
           \vis{\set{\beta}}\!,\Pred,\Post
				 } {s \OTWeakarrow {\alpha} s'} \in \WT
}\qquad \WTDeux
%\inferrule{
% \openrule
%         {
%           \set{\beta},\Pred,\Post}
%         {s \OTarrow {\tau} s'} \in \mathcal{T}
%\\
% \openrule
%         {
%           \set{\gamma},\Pred\,',({x_k\gets e_k})^{k\in K}   }
%         {s' \OTWeakarrow {\tau} s''} \in \WT
%}
%{ \openrule
%         {
%           \vis{\set{\beta}}\dotcup\set{\gamma},\Pred\land\Pred\,'\subst{\Post\,},
%				({x_k\gets (e_k\subst{\Post\,})})^{k\in K} } 
%         {s \OTWeakarrow {\tau} s''} \in \WT
%}
 \end{mathpar}
 and
%\noindent{\bf Other transitions:}
\begin{mathpar}
  \mprset {vskip=.5ex}
\inferrule {\openrule
         {
           \set{\gamma_1},\Pred_1,\Post_1   }
         {s \OTWeakarrow {\tau} s_1} \in \WT
\qquad
\openrule
         {
           \set{\gamma_2},\Pred_2,\Post_2  }
         {s_1 \OTWeakarrow {\alpha} s_2} \in \WT
\qquad
\openrule
         {
           \set{\gamma_3},\Pred_3,({x_k\gets e_k})^{k\in K}    }
         {s_2 \OTWeakarrow {\tau} s'} \in\WT
\\
\Pred=\Pred_1\land\Pred_2\subst{\Post_1}\land (\Pred_3\subst{\Post_2})\subst{\Post_1}
}
{
\openrule
         {
           \set{\gamma_1}\dotcup \set{\gamma_2}\dotcup\set{\gamma_3},
		\Pred,
				({x_k\gets (e_k\subst {\Post_2})\subst{\Post_1}})^{k\in K} } 
         {s \OTWeakarrow {\alpha} s'} \in\WT
} \WTTrois
\end{mathpar}

@@

\subsection{Weak bisimulation for open pNets}
\begin{definition}[Non-observability of silent actions for pNets]~\\
A pNet $\mylangle \pNet_i^{i\in I} , \Sort_j^{j\in J}, \set{\symb{SV}}\myrangle$
 \emph{cannot observe silent actions} if it verifies:\\ $\forall i\in I\uplus J.\, \SV{(i\mapsto \tau)}{\tau}{\True}\in \set{\symb{SV}}$ and $\forall SV_k\in \set{\symb{SV}}, \forall i,j\in I_k\uplus J_k.\,i\neq j \Rightarrow \alpha'_k\neq\tau$
\end{definition}

@@

%%    \left( fresh \ \set{a_i}, \set{b_j}, v_x.\ \

@@

       	\begin{mathpar} 
       	%    \left( fresh \ \set{a_i}, \set{b_j}, v_x.\ \
       	\openrule
       	{
       		\beta_{j p x}^{j\in J_{px}}, \Pred_{OT_{px}},\Post_{OT_{px}}}
       	{t_p \OTarrow {\alpha_{px}} t_{p x}}\qquad (*)
       	%         \right)
       	\end{mathpar}
